What's hot and what's not:\\Windowing developer topic analysis

### 2009-04-11 lda-paper.tex
Edits and figure clear up

Imagine that a development team had agreed upon a set of features to
implement for a development iteration, but these features were not
finished by the end of the iteration. The developers insist they
focussed on those features, yet their manager is not sure what
happened. In a scenario like this, topic analysis can help tease out
the topics of development that the developers were focussed on. Topic
analysis could highlight the orthogonal issues that developers were
spending time on.  Topic analysis extracts groups of words from
documents (commit log messages) that are word distributions, topics,
which characterize clusters of documents. These extracted topics often
correlate with actual development topics that the developers
discussed. By applying topic analysis locally to an iteration, both
the developers and managers might be able to see what were the topics
of development during that iteration. %XXX RM during that iteration

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS commit comments made when the user
commits changes to files in a repository. A \emph{word distribution}
is the summary of a message by word count. Each word distribution will
be over the words in all messages. However, most words will not appear
in a message, a word distribution is effectively a word count divided
by the message size. A \emph{topic} is a word distribution. It is
essentially a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is a one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

@@

Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help bug
localization by querying for documents that are related to a bug's topic. They
used LDA to build a topic model and then queried against it using
sample documents. These queries would indicate the proportion of
topics that were associated with the query and thus possible the
related documents.

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
%location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
location~\cite{1374321,10.1109/ICPC.2007.13}
Concept analysis aims to automatically extract concepts from source
code or documents that are related to the source code.  Concept location concerns how
high-level concepts relate to low level entities such as source code. For
example, when fixing a bug how and where do the concepts in the bug
report map back to the source code?  Semantic Clustering has also been
used for similar purposes~\cite{1698774,1566153} as it is similar to
LSI.

@@

Our technique differs from the previous techniques because we apply
LDA locally to month-long windows of commit comments, whereas other
approaches apply LDA once to the entire project. This allows us to
examine the time-based windowing of topics over its development
history.

@@

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
 examined the top 10 most frequent words in that topic.  We
found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
be viewed as stop words. There were notable transitional topics such
as in the first window the word \emph{BitKeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of Bitkeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with only a little familiarity with the code base that
naming the topic was straightforward.

A sampling of the notable topic words is displayed in Table
\ref{tab:portability}, we chose topics that we felt confident we could
name.  To name each topic we selected a term that seemed to best
summarized that topic.  After extracting these topics, we attempted to
track the evolution of topics by visualizing the topics and joining
similar topics into trends.  Figure \ref{fig:lda} displays a sample
plot of the extracted topics in Table \ref{tab:portability}.

@@

2001 &  Aug &      TABLES row version \\
\hline         
2001 &  Sep &      update checksum merge \\
% 2001 &  Oct &      fixed fix \\
% 2001 &  Dec &      HPUX SCO fix \\
% \hline         
% 2002 &  Feb &      net buffer length  max\_allowed\_packet \\
% 2002 &  Mar &      small buf fix  \\
% \hline         
% 2002 &  May &      [popular] fix SCO OSF1 table\_name \\
% 2002 &  Nov &      HPUX11 compiler HP \\
% \hline         
% 2003 &  Feb &      Linux errno   \\
% 2003 &  Mar &      alarm bookmark bug \\
% \hline         
% 2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Sampled topics from MySQL 3.23, some with continuous topics. These tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit logs.}
\label{tab:portability}
\end{table}

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. This is the
    kind of plot we eventually want to illustrate: named topics and
    topic trends. The horizontal axis is time by month examined. The
    vertical axis is used to stack topics that occur at the same
    time. Longer topics are topics which recur in adjacent
    window. Colors are arbitrary.}
  \label{fig:lda}
\end{figure*}

@@

Our methodology is to first extract the commits from SCS repositories
such as CVS or Bitkeeper. Then we extract the revisions and commits,
grouping the revisions into commits if necessary. We extract the
commit log message of each commit, count the occurrence of each word
in the message, remove stop words, and then produce word distributions
for each message.

@@

The documents we are analyzing are the commit log comments i.e., the
comments that are added when revisions to the project are committed.
Documents are converted into word count distributions. These
distributions are then normalized by the size of the message. After
all messages are processed the distributions are extended to include
all words within all of the  distributions.

@@

We then cluster these topics into trends by comparing them to each
other.  Each topic is compared to each other topic in the system,
given a certain threshold of top 10 similarity, such as 8 out of 10
words. Then given this topic similarity matrix, we find the transitive
closure of similar topics; that is, if we model topics are nodes and
arcs as similarity, we fill flood along the similarity arcs until we
have partitioned the topics into clusters of similar topics. Figure
\ref{fig:closure} illustrates clustering of topics by the
topic similarity. These clusters of topics are called trends. A trend
is probably interesting if it contains more than one topic.

This technique has weaknesses in that nodes that are a few neighbors
away in similarity might not share any similar words.  We do this
because we want to be able to color or highlight topics that are
related to each other and track them throughout time.

@@

\begin{figure*}
  \centering
  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
  \caption{Compact-trend-view shows topics per month for MaxDB
    7.500. The x-axis is time in months, the Y axis is used to stack
    topics occurring at the same time. Trends that are continuous are
    plotted as continuous blocks. The top 10 words in the topics are
    joined and embedded in box representing the topics.}
  \label{fig:topicsmear}
\end{figure*}

>>>>

Imagine that a development team had agreed upon a set of features to
implement for a development iteration, but these features were not
finished by the end of the iteration. The developers insist they
focussed on those features, yet their manager is not sure what
happened. In a scenario like this, topic analysis can help tease out
the topics of development that the developers were focussed on. Topic
analysis could highlight the orthogonal issues that developers were
spending time on.  Topic analysis extracts groups of words from
documents (commit log comments) that are word distributions, topics,
which characterize clusters of documents. These extracted topics often
correlate with actual development topics that the developers
discussed. By applying topic analysis locally to an iteration, both
the developers and managers might be able to see what were the topics
of development during that iteration. %XXX RM during that iteration

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS commit log comments made when the user
commits changes to files in a repository. A \emph{word distribution}
is the summary of a message by word count. Each word distribution will
be over the words in all messages. However, most words will not appear
in a message, a word distribution is effectively a word count divided
by the message size. A \emph{topic} is a word distribution, a
set of words that form a word distribution that is unique and
independent within the set of documents in our total corpus. One could
think of a topic as a distribution of the centroid of a group of
messages. In this paper we often summarize topics by the top 10 most
frequent words of their word distribution.  A \emph{trend} is a one or
more similar topics that recur over time.  Trends are particularly
important, as they indicate long-lived and recurring topics that may
provide key insights into the development of the project.

@@

Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help bug
localization by querying for documents that are related to a bug's topic. They
used LDA to build a topic model and then queried against it using
sample documents. These queries would indicate the proportion of
topics that were associated with the query and the
related documents.

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
%location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
location~\cite{1374321,10.1109/ICPC.2007.13}
Concept analysis aims to automatically extract concepts from source
code or documents that are related to the source code.  Concept location concerns how
high-level concepts relate to low level entities such as source code. For
example, when fixing a bug, how and where do the concepts in the bug
report map back to the source code?  Semantic Clustering has also been
used for similar purposes~\cite{1698774,1566153} as it is similar to
LSI.

@@

Our technique differs from the previous techniques because we apply
LDA locally to month-long windows of commit log comments, whereas other
approaches apply LDA once to the entire project. This allows us to
examine the time-based windowing of topics over its development
history.

@@

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
examined the top 10 most frequent words in that topic.  We found there
were common words across topic clusters, such as \emph{diffs},
\emph{annotate} and \emph{history}; these words probably should be
viewed as stop words. There were notable transitional topics such as
in the first window the word \emph{BitKeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of BitKeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with only a little familiarity with the code base that
naming the topic was straightforward.

A sampling of the notable topic words is displayed in Table
\ref{tab:portability}, we chose topics that we felt confident we could
name.  To name each topic we selected a term that seemed to best
summarized that topic.  After extracting these topics, we attempted to
track the evolution of topics by visualizing the topics and joining
similar topics into trends.  Figure \ref{fig:lda} displays a manually
created sample plot of the extracted topics in Table
\ref{tab:portability}.

@@

2001 &  Aug &      TABLES row version \\
\hline         
2001 &  Sep &      update checksum merge \\
% 2001 &  Oct &      fixed fix \\
% 2001 &  Dec &      HPUX SCO fix \\
% \hline         
% 2002 &  Feb &      net buffer length  max\_allowed\_packet \\
% 2002 &  Mar &      small buf fix  \\
% \hline         
% 2002 &  May &      [popular] fix SCO OSF1 table\_name \\
% 2002 &  Nov &      HPUX11 compiler HP \\
% \hline         
% 2003 &  Feb &      Linux errno   \\
% 2003 &  Mar &      alarm bookmark bug \\
% \hline         
% 2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Sampled topics from MySQL 3.23, some with continuous topics. These tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit comments.}
\label{tab:portability}
\end{table}

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. This is the
    kind of plot we eventually want to illustrate: named topics and
    topic trends. The horizontal axis is time by month examined. The
    vertical axis is used to stack topics that occur at the same
    time. Longer topics are topics which recur in adjacent
    windows. Colors are arbitrary.}
  \label{fig:lda}
\end{figure*}

@@

Our methodology is to first extract the commits from SCS repositories
such as CVS or Bitkeeper. Then we extract the revisions and commits,
grouping the revisions into commits if necessary. We extract the
commit log comment of each commit, count the occurrence of each word
in the message, remove stop words, and then produce word distributions
for each message.

@@

The documents we are analyzing are the commit log comments i.e., the
comments that are added when revisions to the project are committed.
Documents are converted into word count distributions. These
distributions are then normalized by the size of the message. After
all messages are processed the distributions are extended to include
all words from all of the  distributions.

@@

We then cluster these topics into trends by comparing them to each
other.  Each topic is compared to each other topic in the system,
given a certain threshold of top 10 similarity, such as 8 out of 10
words. Then given this topic similarity matrix, we find the transitive
closure of similar topics; that is, if we model topics are nodes and
arcs as similarity, we fill flood along the similarity arcs until we
have partitioned the topics into clusters of similar topics. Figure
\ref{fig:closure} illustrates clustering of topics by 
topic similarity. These clusters of topics are called trends. A trend
is probably interesting if it contains more than one topic.

This technique has weaknesses in that nodes that are a few neighbors
away in similarity might not share any similar words.  We use this measure
because we want to be able to color or highlight topics that are
related to each other and track them throughout time.

@@

\begin{figure*}
  \centering
%  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
  \includegraphics[width=1.0\textwidth]{time-smear-plot}%fixed-time-smear-plot-scaled}
  \caption{Compact-trend-view shows topics per month for MaxDB
    7.500. The x-axis is time in months, the Y axis is used to stack
    topics occurring at the same time. Trends that are continuous are
    plotted as continuous blocks. The top 10 words in the topics are
    joined and embedded in box representing the topics.}
  \label{fig:topicsmear}
\end{figure*}
